---
title: "Assignment 5"
author: "Christina Treacy and Hayden Waters"
date: "May 29, 2026"
format: pdf
editor: visual
---

First, we install the relevant packages.

```{r}
library(tidyverse)
library(here)
library(usethis)
library(broom)
```

**Fire Risk Program Contract** 

Inputs Outputs Parameters

# -----------------------------

# Inputs & Parameters: Set Up the Data Frame

# -----------------------------

```{r}
# Set seed so the random numbers stay the same each time
set.seed(123)

# Create 100 site IDs
site_id <- 1:100

# Create made-up environmental data for each site
temperature <- round(rnorm(100, mean = 31, sd = 5), 1)       # degrees Celsius
humidity <- round(runif(100, min = 10, max = 55), 1)         # percent
wind_speed <- round(runif(100, min = 5, max = 55), 1)        # km/hr
fuel_moisture <- round(runif(100, min = 3, max = 18), 1)     # percent
slope <- round(runif(100, min = 0, max = 45), 1)             # degrees
distance <- round(runif(100, min = 0, max = 5000), 0)        # meters

# Combine all variables into one data frame
fire_sites <- data.frame(
  site_id,
  temperature,
  humidity,
  wind_speed,
  fuel_moisture,
  slope,
  distance
)

# View the data frame & save as a CSV
fire_sites

write.csv(fire_sites, "fire_sites.csv")
```

# -----------------------------

# Module 1: Climate conditions

# -----------------------------

*Module 1 Contract*
This function combines temperature, humidity, and wind speed and divides each variable by a rough upper value to put them on a similar scale. For the temperature calculation, we assume that around 40°C is a very hot fire-weather temperature, so higher temperatures will increase this factor. For the humidity calculation, air with lower humidity (drier air) signifies higher fire risk. Wind speed is divided by 60 because we assume that 60 km/hr is a strong wind value for this model, and higher wind speeds increase fire risk.Essentially, this function says: hot + dry + windy = higher climate risk.

Inputs: average temperature (degrees Celsius), humidity (percent), and wind speed (km/hr) at each location
Output: Climate risk factor
Parameters: hot fire-weather temperature (ht) = 40°C; strong wind value (wv) = 60 km/hr is a strong wind value
What the function will do: Compute the level of wildfire risk based on climate conditions alone for each site using the following equation:
climate_risk = (temperature / ht) +  ((100 - humidity) / 100) +  (wind_speed / wv)

*Call Module 1 Function*
```{r}
#Call climate_conditions function
source(here("climate_conditions.R"))
```

# -----------------------------

# Module 2: Land conditions

# ----------------------------

Inputs:
Fuel moisture: The percentage of moisture by weight contained by living and dead vegetation averaged across a site. 
Slope: The average topographic slope of each location in degrees. 

Output: 
Land Fire Probability: The probability of a fire igniting and spreading based purely on land conditions.

*Call Module 2 Function*
```{r}
#Call land_conditions function
source(here("land_conditions.R"))
```

# -----------------------------

# Module 3: Wildfire likelihood

# -----------------------------

*Module 3 Contract*

*Call Module 3 Function and Run For Loop*
```{r}
#Call wildfire_likelihood function
source(here("wildfire_likelihood.R"))
source("wildfire_likelihood.R")

# After creating fire_sites, run the Module 3 loop
fire_sites$climate_fire_probability <- NA
fire_sites$land_fire_probability <- NA
fire_sites$wildfire_probability <- NA

for (i in 1:nrow(fire_sites)) {
  
  fire_sites$climate_fire_probability[i] <- climate_conditions(
    fire_sites$temperature[i],
    fire_sites$humidity[i],
    fire_sites$wind_speed[i]
  )
  
  fire_sites$land_fire_probability[i] <- land_conditions(
    fire_sites$fuel_moisture[i],
    fire_sites$slope[i]
  )
  
  fire_sites$wildfire_probability[i] <- wildfire_likelihood(
    fire_sites$climate_fire_probability[i],
    fire_sites$land_fire_probability[i]
  )
}
```

# -----------------------------

# Module 4: Final risk factor

# -----------------------------

*Module 4 Contract*

*Module 4 Calculation*
```{r}
# Step 1: Calculate log10 risk scores
fire_sites$risk_score <- NA

for (i in 1:nrow(fire_sites)) {
  
  raw_risk <- risk_factor(
    fire_sites$wildfire_probability[i],
    fire_sites$distance[i]
  )
  
  fire_sites$risk_score[i] <- log10(raw_risk + 0.001)
}


# Step 2: Calculate cutoffs based on the log10 risk scores
low_cutoff <- quantile(fire_sites$risk_score, 0.33, na.rm = TRUE)
high_cutoff <- quantile(fire_sites$risk_score, 0.66, na.rm = TRUE)


# Step 3: Classify risk categories
fire_sites$risk_category <- NA

for (i in 1:nrow(fire_sites)) {
  
  if (is.na(fire_sites$risk_score[i])) {
    fire_sites$risk_category[i] <- NA
    
  } else if (fire_sites$risk_score[i] < low_cutoff) {
    fire_sites$risk_category[i] <- "Low"
    
  } else if (fire_sites$risk_score[i] < high_cutoff) {
    fire_sites$risk_category[i] <- "Medium"
    
  } else {
    fire_sites$risk_category[i] <- "High"
  }
}


# Step 4: Check results
table(fire_sites$risk_category)
summary(fire_sites$risk_score)
```
```{r}
#Call risk_factor function
source(here("risk_factor_function.R"))

fire_sites$risk_score <- NA

for (i in 1:nrow(fire_sites)) {
  
  raw_risk <- risk_factor(
    fire_sites$wildfire_probability[i],
    fire_sites$distance[i]
  )
  
  fire_sites$risk_score[i] <- log10(raw_risk + 0.001)
}

# Step 2: Calculate cutoffs based on the log10 risk scores
low_cutoff <- quantile(fire_sites$risk_score, 0.33, na.rm = TRUE)
high_cutoff <- quantile(fire_sites$risk_score, 0.66, na.rm = TRUE)


# Step 3: Classify risk categories
fire_sites$risk_category <- NA

for (i in 1:nrow(fire_sites)) {
  
  if (is.na(fire_sites$risk_score[i])) {
    fire_sites$risk_category[i] <- NA
    
  } else if (fire_sites$risk_score[i] < low_cutoff) {
    fire_sites$risk_category[i] <- "Low"
    
  } else if (fire_sites$risk_score[i] < high_cutoff) {
    fire_sites$risk_category[i] <- "Medium"
    
  } else {
    fire_sites$risk_category[i] <- "High"
  }
}


# Step 4: Check results
table(fire_sites$risk_category)
summary(fire_sites$risk_score)

#Classify the sites into low, medium, and high risk:

fire_sites$risk_category <- NA

for (i in 1:nrow(fire_sites)) {
  
  if (fire_sites$risk_score[i] < 0.5) {
    fire_sites$risk_category[i] <- "Low"
  } else if (fire_sites$risk_score[i] < 1.2) {
    fire_sites$risk_category[i] <- "Medium"
  } else {
    fire_sites$risk_category[i] <- "High"
  }
}
```

# -----------------------------

# Final Output: Run For Loop to Classify Risk Categories

# -----------------------------

```{r}

for (i in 1:nrow(fire_sites)) {

 fire_sites$climate_risk_factor[i] <- climate_conditions(
   fire_sites$temperature[i],
   fire_sites$humidity[i],
   fire_sites$wind_speed[i]
 )

 fire_sites$ignition_probability[i] <- land_conditions(
   fire_sites$fuel_moisture[i],
   fire_sites$slope[i]
 )

 fire_sites$wildfire_probability[i] <- wildfire_likelihood(
   fire_sites$climate_risk_factor[i],
   fire_sites$ignition_probability[i]
 )

 fire_sites$risk_score[i] <- risk_factor(
   fire_sites$wildfire_probability[i],
   fire_sites$distance[i]
 )
}

# Inspect the finished data frame:
head(fire_sites)

summary(fire_sites$risk_score)

#Classify the sites into low, medium, and high risk:

fire_sites$risk_category <- NA

for (i in 1:nrow(fire_sites)) {

 if (fire_sites$risk_score[i] < 1.5) {
   fire_sites$risk_category[i] <- "Low"
 } else if (fire_sites$risk_score[i] < 2.3) {
   fire_sites$risk_category[i] <- "Medium"
 } else {
   fire_sites$risk_category[i] <- "High"
 }
}

```
