---
title: "Assignment 5"
author: "Christina Treacy and Hayden Waters"
date: "May 29, 2026"
format: pdf
editor: visual
---

First, we install the relevant packages.

```{r}
library(tidyverse)
library(here)
library(usethis)
library(broom)
```

**Fire Risk Program Contract** 

Inputs Outputs Parameters

# -----------------------------

# Inputs & Parameters: Set Up the Data Frame

# -----------------------------

```{r}
# Set seed so the random numbers stay the same each time
set.seed(123)

# Create 100 site IDs
site_id <- 1:100

# Create made-up environmental data for each site
temperature <- round(rnorm(100, mean = 31, sd = 5), 1)       # degrees Celsius
humidity <- round(runif(100, min = 10, max = 55), 1)         # percent
wind_speed <- round(runif(100, min = 5, max = 55), 1)        # km/hr
fuel_moisture <- round(runif(100, min = 3, max = 18), 1)     # percent
slope <- round(runif(100, min = 0, max = 45), 1)             # degrees
distance <- round(runif(100, min = 0, max = 5000), 0)        # meters

# Combine all variables into one data frame
fire_sites <- data.frame(
  site_id,
  temperature,
  humidity,
  wind_speed,
  fuel_moisture,
  slope,
  distance
)

# View the data frame & save as a CSV
fire_sites

write.csv(fire_sites, "fire_sites.csv")
```

# -----------------------------

# Module 1: Climate conditions

# -----------------------------

*Module 1 Contract*

*Module 1 Calculation*
```{r}
#Call climate_conditions function

```

# -----------------------------

# Module 2: Land conditions

# ----------------------------

Inputs:
Fuel moisture: The percentage of moisture by weight contained by living and dead vegetation averaged across a site. 
Slope: The average topographic slope of each location in degrees. 

Output: 
Land Fire Probability: The probability of a fire igniting and spreading based purely on land conditions. 
*Module 2 Calculation*
```{r}
#Call land_conditions function
source(here("land_conditions.R"))
```

# -----------------------------

# Module 3: Wildfire likelihood

# -----------------------------

*Module 3 Contract*

*Module 3 Calculation*
```{r}
#Call wildfire_likelihood function
source(here("wildfire_likelihood.R"))
```

# -----------------------------

# Module 4: Final risk factor

# -----------------------------

*Module 4 Contract*

*Module 4 Calculation*
```{r}
#Call risk_factor function

```

# -----------------------------

# Final Output: Run For Loop to Classify Risk Categories

# -----------------------------

```{r}
for (i in 1:nrow(fire_sites)) {

 fire_sites$climate_risk_factor[i] <- climate_conditions(
   fire_sites$temperature[i],
   fire_sites$humidity[i],
   fire_sites$wind_speed[i]
 )

 fire_sites$ignition_probability[i] <- land_conditions(
   fire_sites$fuel_moisture[i],
   fire_sites$slope[i]
 )

 fire_sites$wildfire_probability[i] <- wildfire_likelihood(
   fire_sites$climate_risk_factor[i],
   fire_sites$ignition_probability[i]
 )

 fire_sites$risk_score[i] <- risk_factor(
   fire_sites$wildfire_probability[i],
   fire_sites$distance[i]
 )
}

# Inspect the finished data frame:
head(fire_sites)

summary(fire_sites$risk_score)

#Classify the sites into low, medium, and high risk:

fire_sites$risk_category <- NA

for (i in 1:nrow(fire_sites)) {

 if (fire_sites$risk_score[i] < 1.5) {
   fire_sites$risk_category[i] <- "Low"
 } else if (fire_sites$risk_score[i] < 2.3) {
   fire_sites$risk_category[i] <- "Medium"
 } else {
   fire_sites$risk_category[i] <- "High"
 }
}

```
